ShrinkNets

### 2018-01-05 paper/paper.tex

When designing Neural Networks, finding the appropriate size (width and depth)
is key. In particular, these hyper parameters have a strong correlation with
over/underfitting \srm{give reference} \gl{I did some litterature review and
  actually this statement is quite controversial, should we rephrase ?}. The
main problem is that we have no reliable way to find them [ref]. Decades of
experimentation led to some heuristics [ref] that try to prune that immense
space of possible network sizes.  For example, researchers have used strategies
such as random search \gl{Should we also move the references here if we remove
  the pargraph from the related work ?}, meta-gradient descent and parser trees
to reduce find good parameters without exhaustively exploring the entire
hyper-parameter space. Although these strategies help with finding a good set
of hyper-parameters, they still require a compute-intensive search of the
space.

@@

This choice come from the Lasso loss [ref] and we use it for the same reason:
it introduces sparsity. The bigger the $\lambda$ the more entries in $\theta$
are set to zer0, and since zero entries in $\theta$ correspond to dead neurons,
lambda effectively control the number of neurons/channels in the entire
network.
Next, we explain how to implement ShrinkNets efficiently.

\textbf{Software architecture\footnote{The code is available as Python/PyTorch
    library on \url{http://github.com/mitdbg/fastdeepnets} \gl{Should we rename
      the repository ?}}: }\textit{Filter Vectors} can easily be implemented in
a few lines of code in many existing deep learning frameworks. However, in
ShrinkNets, we assume that we start with obviously oversized networks. If
disabled neurons are not quickly removed, the overhead might cause the training
process to be significantly slower than classic neural networks. This is why we
want to support what we call garbage collection: removing the weights in the
layers that are responsible (or using) features that are disabled.  \par To be
able to catpure this relationship between layers, we agumented \textit{Pytorch}
(the framework we used as the foundation of our implementation) with a graph
structure very similar to the one available in Keras [ref].  In this graph,
edges are effectively event hubs responsible for propagating \textit{feature
  removal} events to its endpoints. ShrinkNet layers are designed to emit and
react to these events but also propagate the event further if has an impact at
the other endpoint of the layer (input or output) \gl{What I mean here is: if
  the event comes from the output, it might be propagated to the inputs, and
  the other way around}.  \par This event-based implementation coordinated by
edges makes it very easy to integrate new layers in the library. We even
provide automatic wrapping for layers from \textit{PyTorch} if they have no
internal state. For more complex ones, we provide utilitaries that makes the
implementation very concise.

@@

To demonstrate that the approach is viable, we will first show that Shrinking
Networks converge properly. For this experiment we trained a one hidden layer
neural network with one filter layer to control the number of hidden units. We
initialized the models with $10000$ neurons and trained them on \texttt{MNIST}
using different regularization factors ($\lambda$). We summarized the results
in \autoref{convergence_plot}. We can see on this plot is that
for a given $\lambda$ the number of hidden units seems to always converge to
some value. It has two implications: it seems that $\lambda$ is, as we would think,
a proxy of the network size (bigger $\lambda$ imply smaller networks). Secondly,
it indicates that the spikes we see in the regular spikes we see in the loss that occur when neurons are disabled will eventually disapear because the number of neurons reach a plateau. \gl{I feel we should give a conclusion for this paragraph but
I am not sure what to say}
\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{convergence}
\caption{Evolution of the number of hidden units and loss over time on the \texttt{MNIST} dataset for different $\lambda$ values \label{convergence_plot}}
\end{center}
\end{figure}

@@

\par To isolate the effect of ShrinkNets we will consider two very simple
architectures: multi-layer perceptrons (with three hidden layers) and the
\textit{LeNet-5} model. For this experiment we assume we have no prior
information about the size except an upper bound of $50$ channels per
convolutional layer and $5000$ neuron for fully connected layers.  We try to
find the appropriate size of the network using classical neural networks and
ShrinkNets. To make sure that we are as fair as possible and since one might
argue that ShrinkNets is also regularizing the network in addition to finding
the size, we will also considered regularized classic networks using the most
commonly used technique: the $L_2$ penalty [ref].
\par The experiment consist in sampling parameters randomly for each class of
network, train them for the same amount of epochs (100 for MLPs and 200 for CNNs). Pick the epoch that performed the best on the validation set and evaluate it
on the testing set. We repeat the process for 50 models. This evaluation was done
on \texttt{MNIST} [ref], \texttt{FashionMNIST} [ref] and \texttt{CIFAR10} [ref]. The distribution of the testing accuracy we obtained is summarized on \autoref{hyper_opt_res}.

@@

\par There are many methods in the litterature that try to simplify the neural
network structure. Most of them focuses on removing connections (eg: [ref],
[ref]).  However, most hardware (GPU's, TPUS, TensorCores) do not really
leverage sparsity in the connections efficiently. On the other side, ShrinkNets
and some others [ref] and [ref] try to remove entire neuron instead of
connections. This is very usefull because it reduces the size of the matrices,
producing speed-up on every device. ShrinkNets improves on [ref] because it
removes neurons during training, speeding up the rest of the process and beats
[ref] on convergence speed and because it does not need to change the
optimizer. To the best of our knowledge this is also the first contribution
that tries to learn the number of channels of a convolutional neural network.

@@

In this paper we presented a novel technique to guess a resonable network size
based on single parameter $\lambda$ that control the tradeof between loss and
the size. We demonstrated that it works both on fully connected networks and
convolutional neural networks.  Even though the firsts results seem promising,
there are many ways we could improve. In the current
implementation we only "learn" the number of features (neurons or channels). We
could try to augment it with dynamic number of layers as seen in [ref] to be
able to determine the entire architecture. \par We saw on Figure
\ref{convergence_plot} that the loss temporarily suffers from the removal of
neurons. It is likely that the loss would be more stable if the number of
neurons converged faster or neurons disappeared slower. For this reason we plan
to explore proximal gradient methods to optimize the filter vectors and/or
randomize neuron removals. \par During our evaluation we picked small datasets
mainly to be able to train many models and have statistically significant
distributions. With more computation resources and time, we could see if it
generalizes to bigger datasets and other architectures like \texttt{ResNet}
[ref] (small modifications to the existing code base are required to support
them)

\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-bibliography}

>>>>

When designing Neural Networks, finding the appropriate size (width and depth)
is key. In particular, these hyper parameters have a strong correlation with
over/underfitting \srm{give reference} \gl{I did some litterature review and
  actually this statement is quite controversial, should we rephrase ?}. We
have no reliable way to find them but decades of experimentation led to some
heuristics ~\cite{Bengio2012a} that try to prune that immense space of possible
network sizes and suggest range of values to explore. To select the best
candidates , researchers have used strategies such as random search \gl{Should
  we also move the references here if we remove the pargraph from the related
  work ?}, meta-gradient descent ~\cite{Pedregosa2016} and Parzen Estimators
~\cite{Bergstra2011a} to reduce find good parameters without exhaustively
exploring the entire hyper-parameter space. Although these strategies help with
finding a good set of hyper-parameters, they still require a compute-intensive
search of the space.

@@

This choice come from the Lasso loss ~\cite{Tibshirani1996} and we use it for the same reason:
it introduces sparsity. The bigger the $\lambda$ the more entries in $\theta$
are set to zer0, and since zero entries in $\theta$ correspond to dead neurons,
lambda effectively control the number of neurons/channels in the entire
network.
Next, we explain how to implement ShrinkNets efficiently.

\textbf{Software architecture\footnote{The code is available as Python/PyTorch
    library on \url{http://github.com/mitdbg/fastdeepnets} \gl{Should we rename
      the repository ?}}: }\textit{Filter Vectors} can easily be implemented in
a few lines of code in many existing deep learning frameworks. However, in
ShrinkNets, we assume that we start with obviously oversized networks. If
disabled neurons are not quickly removed, the overhead might cause the training
process to be significantly slower than classic neural networks. This is why we
want to support what we call garbage collection: removing the weights in the
layers that are responsible (or using) features that are disabled.  \par To be
able to catpure this relationship between layers, we agumented \textit{Pytorch}
~\cite{paszke2017automatic} (the framework we used as the
foundation of our implementation) with a graph structure very similar to the
one available in Keras ~\cite{chollet2015keras}.  In this graph, edges are
effectively event hubs responsible for propagating \textit{feature removal}
events to its endpoints. ShrinkNet layers are designed to emit and react to
these events but also propagate the event further if has an impact at the other
endpoint of the layer (input or output) \gl{What I mean here is: if the event
  comes from the output, it might be propagated to the inputs, and the other
  way around}.  \par This event-based implementation coordinated by edges makes
it very easy to integrate new layers in the library.  We even provide automatic
wrapping for layers from \textit{PyTorch} if they have no internal state. For
more complex ones, we provide utilitaries that
makes the implementation very concise.

@@

To demonstrate that the approach is viable, we will first show that Shrinking
Networks converge properly. For this experiment we trained a one hidden layer
neural network with one filter layer to control the number of hidden units. We
initialized the models with $10000$ neurons and trained them on \texttt{MNIST}
~\cite{Lecun1998} using different regularization factors ($\lambda$). We
summarized the results in \autoref{convergence_plot}. We can see on this plot
is that for a given $\lambda$ the number of hidden units seems to always
converge to some value. It has two implications: it seems that $\lambda$ is, as
we would think, a proxy of the network size (bigger $\lambda$ imply smaller
networks). Secondly, it indicates that the spikes we see in the regular spikes
we see in the loss that occur when neurons are disabled will eventually
disapear because the number of neurons reach a plateau. \gl{I feel we should
  give a conclusion for this paragraph but I am not sure what to say}
\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{convergence}
\caption{Evolution of the number of hidden units and loss over time on the \texttt{MNIST} dataset for different $\lambda$ values \label{convergence_plot}}
\end{center}
\end{figure}

@@

\par To isolate the effect of ShrinkNets we will consider two very simple
architectures: multi-layer perceptrons (with three hidden layers) and the
\textit{LeNet-5} model ~\cite{Lecun1998}. For this experiment we assume we have no prior
information about the size except an upper bound of $50$ channels per
convolutional layer and $5000$ neuron for fully connected layers.  We try to
find the appropriate size of the network using classical neural networks and
ShrinkNets. To make sure that we are as fair as possible and since one might
argue that ShrinkNets is also regularizing the network in addition to finding
the size, we will also considered regularized classic networks using the most
commonly used technique: the $L_2$ penalty ~\cite{Ng2004}.
\par The experiment consist in sampling parameters randomly for each class of
network, train them for the same amount of epochs (100 for MLPs and 200 for CNNs). Pick the epoch that performed the best on the validation set and evaluate it
on the testing set. We repeat the process for 50 models. This evaluation was done
on \texttt{MNIST} ~\cite{Lecun1998}, \texttt{FashionMNIST} ~\cite{Xiao2017} and \texttt{CIFAR10} ~\cite{Krizhevsky2009}. The distribution of the testing accuracy we obtained is summarized on \autoref{hyper_opt_res}.

@@

\par There are many methods in the litterature that try to simplify the neural
network structure. Most of them focuses on removing connections (eg: ~\cite{Cun},
~\cite{Han2015}).  However, most hardware (GPU's, TPUS ~\cite{Jouppi2017}, TensorCores) do not really
leverage sparsity in the connections efficiently (as explained in ~\cite{Han2016}). On the other side, ShrinkNets
and some others ~\cite{Scardapane2017} and ~\cite{Philipp} try to remove entire neuron instead of
connections. This is very usefull because it reduces the size of the matrices,
producing speed-up on every device. ShrinkNets improves on
~\cite{Scardapane2017} because it removes neurons during training, speeding up
the rest of the process and beats ~\cite{Philipp} on convergence speed and because it
does not need to change the optimizer. To the best of our knowledge this is
also the first contribution that tries to learn the number of channels of a
convolutional neural network.

@@

In this paper we presented a novel technique to guess a resonable network size
based on single parameter $\lambda$ that control the tradeof between loss and
the size. We demonstrated that it works both on fully connected networks and
convolutional neural networks.  Even though the firsts results seem promising,
there are many ways we could improve. In the current implementation we only
"learn" the number of features (neurons or channels). We could try to augment
it with dynamic number of layers as seen in ~\cite{meier} to be able to
determine the entire architecture. \par We saw on Figure \ref{convergence_plot}
that the loss temporarily suffers from the removal of neurons. It is likely
that the loss would be more stable if the number of neurons converged faster or
neurons disappeared slower. For this reason we plan to explore proximal
gradient methods to optimize the filter vectors and/or randomize neuron
removals. \par During our evaluation we picked small datasets mainly to be able
to train many models and have statistically significant distributions. With
more computation resources and time, we could see if it generalizes to bigger
datasets and other architectures like \texttt{ResNet} ~\cite{He2016} (small
modifications to the existing code base are required to support them)

\bibliographystyle{ACM-Reference-Format}
\bibliography{bib,custom}
